Rapid long-range disynaptic inhibition explains the formation of cortical orientation maps.

### 2017-02-27 paper.tex
more changes

In this study we have shown how recent findings on dependence of neural transmission delays on the
type of pre- and post-synaptic neurons \citep{Ohana2012} can resolve a long-standing question on how short-range inhibition
can support cortical competition and consequently the development of functional cortical topological organization.
Under simplifying assumptions, we have analytically shown how disynaptic inhibition that is as fast as mono-synaptic 
excitation can extend the effective range of inhibitory interactions, in contrast to the recent analytical 
results showing that in the case of equal synaptic delays on all connections the disynaptic inhibition has negligible effects \citep{Muir2014}.\
We have also shown that these findings are applicable to the problem of functional development in primary visual cortex. We have then proceeded 
to show using computational methods that the proposed models are robust to the addition of other well established features of cortical anatomy, commonly 
ignored by similar studies, including the long-range excitatory connections and mutual inhibition among inhibitory neurons. Finally, we have shown 
that the proposed mechanisms are robust to the variations of the exact delay ratio between the mono-synaptic excitation and di-synaptic inhibition. 
Overall, this study represents an important advance in our understanding of how orientation map development can be supported by the cortical neural substrate. These
results generalize to the development of other functional features in the cortex and other cortical competition based mechanisms in general.

@@

The analytical results in this study were obtained only under simplifying assumptions, specifically instantaneous translation of inputs to membrane potential, equal extent of 
excitatory and inhibitory connections and a lack of inhibitory-to-inhibitory interactions. Even though we have shown computationally that these assumptions are not necessary 
for achieving the cortical competition and the consequential orientation map development sought in this study, further analytical work then can circumvent these simplifications and would 
undoubtedly provide deeper understanding of the dynamics of the studied neural system and its dependence on the various parameters. This sentiment underlies the parameter 
explorations presented here, which show that even though the model is robust to changes in the considered parameters, the existence or not of dynamics supporting development of 
orientation maps forms a complex pattern within the explored parameter spaces. Furthermore, the relative computational complexity of the studied models and the extensive set of parameters
involved preclude systematic search across the full parameter space, and we have only explored parameters that we empirically found to have the biggest impact. Finally, one simplifying assumption 
that we have not treated in this study is the lack of direct thalamic input onto inhibitory cells. Since inhibitory cells in cortical layer 4 do receive thalamic input \citep{Binzegger2004} the 
inclusion of external input in the inhibitory population needs to be considered in the future.

>>>>

In this study we have shown how recent findings on dependence of neural transmission delays on the
type of pre- and post-synaptic neurons \citep{Ohana2012} can resolve a long-standing question on how short-range inhibition
can support cortical competition and consequently the development of functional cortical topological organization.
Under simplifying assumptions, we have analytically shown how disynaptic inhibition that is as fast as mono-synaptic 
excitation can extend the effective range of inhibitory interactions, in contrast to the recent analytical 
results showing that in the case of equal synaptic delays on all connections the disynaptic inhibition has negligible effects \citep{Muir2014}.\
We have also shown that these findings are applicable to the problem of functional development in primary visual cortex. We have then proceeded 
to show using computational methods that the proposed models are robust to the addition of other well established features of cortical anatomy, commonly 
ignored by similar studies, including the long-range excitatory connections and mutual inhibition among inhibitory neurons. Finally, we have shown 
that the proposed mechanisms are robust to the variations of the exact delay ratio between the mono-synaptic excitation and di-synaptic inhibition. 
Overall, this study represents an important advance in our understanding of how orientation map development can be supported by the cortical neural substrate.

Using neuron reconstruction data of the recorded neurons, Ohana \etal,\citep{Ohana2012} identified the 
placement of the different pre-post synaptic combinations on the dendrites of the target cell as the likely origin of this 
transmission delay heterogeneity. They found that the excitatory-to-inhibitory synapses were closer to the soma
of the post-synaptic neuron than the excitatory-to-excitatory and inhibitory-to-excitatory synapses. An axonal origin 
for the transmission delay heterogeneity is unlikely, as Ohana \etal,\citep{Ohana2012} found that all the pre-post combinations were, on
average, equally distant from the pre-synaptic cell body. They further supported the dendritic origin of the observed
delay inhomogeneities by showing in a computational model that the observed delay magnitudes can be explained by
these anatomical findings. Overall this indicates a mechanism for the generation of the differences in 
inter-neuron transmission delays that is likely to generalize beyond the primary visual cortex. We thus suggest that the
results presented in this study generalize to the development of other functional features and other 
cortical competition based mechanisms whose origin is outside of primary visual cortex.

@@

The analytical results in this study were obtained only under simplifying assumptions, specifically instantaneous translation of inputs to membrane potential, equal extent of 
excitatory and inhibitory connections and a lack of inhibitory-to-inhibitory interactions. Even though we have shown computationally that these assumptions are not necessary 
for achieving the cortical competition and the consequential orientation map development sought in this study, further analytical work then can circumvent these simplifications would 
undoubtedly provide deeper understanding of the dynamics of the studied neural system and its dependence on the various parameters. This sentiment underlies the parameter 
explorations presented here, which show that even though the model is robust to changes in the considered parameters, the existence or not of dynamics supporting development of 
orientation maps can form a complex pattern within the explored parameter spaces. This is particularly the case for the model variant with inhibitory-to-inhibitory interactions, in which 
(unlike in the other variants) the inhibitory population gains it's own dynamics. Furthermore, the relative computational complexity of the studied models and the extensive set of parameters
involved preclude systematic search across the full parameter space, and we have only explored parameters that we empirically found to have the biggest impact. Finally, one simplifying assumption 
that we have not treated in this study is the lack of direct thalamic input onto inhibitory cells. Since inhibitory cells in cortical layer 4 do receive thalamic input \citep{Binzegger2004} the 
inclusion of external input in the inhibitory population needs to be considered in the future.
